---
title: "Computing Final Exam"
author: "Dongeun Min"
date: "12/20/2020"
output: pdf_document
---

## Settings
```{r setup}
library("Rcpp")

# Compile C++ code and import function into R
current_path = rstudioapi::getActiveDocumentContext()$path
setwd(dirname(current_path))
sourceCpp("Computing_Final_2020324011.cpp")
```

## Question 1.(a)
```{r}
h = function(x){
  value = (cos(50*x)+sin(20*x))^2
  return(value)
}

hprime1 = function(x){
  value = 2 * (cos(50*x)+sin(20*x)) * (-50*sin(50*x)+20*cos(20*x));
  return(value)
}

hprime2 = function(x){
  temp1 = (-50*sin(50*x)+20*cos(20*x))^2;
  temp2 = (cos(50*x)+sin(20*x))*(-2500*cos(50*x)-400*sin(20*x));
  value = 2*(temp1 + temp2);
  return(value)
}

# newton method
newton = function(init, eps){
  x = init;
  i = 0;
  diff = 1.0;
  while(abs(diff) > eps){
    diff = -hprime1(x)/hprime2(x)
    x = x + diff
    i = i + 1
  } 
  return(x)
}
```

```{r}
theta = seq(1e-7, 1-(1e-7), length=500)
lv = theta
for(i in 1:length(theta)){
  lv[i] = h(theta[i])
}
plot(theta, lv, type="l")

start = c(0.2, 0.3, 0.35, 0.38, 0.4, 0.5, 0.56, 0.6, 0.7, 0.8)
x_hat = rep(0, 10)
like = rep(0, 10)
for (i in 1:length(start)){
  x_hat[i] = newton(start[i], 1e-10)
  like[i] = h(x_hat[i])
  points(x_hat[i], h(x_hat[i]), col=i)
}
x_hat
like
```

The maximum value is 3.832544.  
  

## Question 1.(b)
```{r}
x = matrix(0, 2500, 4)  #initial value = 0
hval = matrix(0, 2500, 4)
r = 0.5
for (j in 1:4){
  for (i in 1:2499){
    a_t = max(x[i,j]-r, 0)
    b_t = min(x[i,j]+r, 1)
    u = runif(1, a_t, b_t)
    T_t = 1/log(i)
    ratio = min(exp((h(u)-h(x[i,j]))/T_t), 1)
    prob = runif(1, 0, 1)
    if (prob < ratio){
      x[i+1,j] = u
    } else {
      x[i+1,j] = x[i,j] }
    # update T_t and T_(t+1) ??
  }
}
for (j in 1:4){
  for (i in 1:2500){
    hval[i,j] = h(x[i,j]) }}

# Plotting
theta = seq(1e-7, 1-(1e-7), length=500)
lv = theta
for (i in 1:length(theta)){
  lv[i] = h(theta[i])
}
plot(theta, lv, type="l")
for (i in 1:2500){
  points(x[i,1], hval[i,1], col=2)
  points(x[i,2], hval[i,2], col=3)
  points(x[i,3], hval[i,3], col=4)
  points(x[i,4], hval[i,4], col=5)
}
x[2500,]
hval[2500,]
```

From the simulated annealing algorithm, it is found that the maximum value is 
found when x is around 0.38 or 0.56.  
The maximum value found is 0.83251.


## Question 2.(b)
```{r}
rikz <- read.table("/Users/michelle/coursework/CS/rikz.txt", header = TRUE)
Y = rikz[,1]  #Response Variable Y: Richness
X = rikz[,3]  #Fixed Effect Variable X: NAP
Z = rikz[,4]  #Random Effect Variable Z: Beach
N = nrow(rikz)
beta = 1  #initial value of beta
D = 0.5  #initial value of D (= variance of b_i)
sig = 0.5  #initial value of sigma(small)

evec = rep(0,3*N)
estep = function(x, y, z, b, d, s){
  exp_s = rep(0,N)
  exp_e = rep(0,N)
  exp_u = rep(0,N)
  for (i in 1:nrow(rikz)) {
    sigma = d*(z[i]^2) + s^2
    exp_s[i] = (d - (d^2)*(z[i]^2)/sigma) + (d*z[i]*(y[i]-x[i]*b)/sigma)^2
    exp_e[i] = (s^2 - (s^2)/sigma) + (s*(y[i]-x[i]*b)/sigma)^2
    exp_u[i] = d*z[i]*(y[i]-x[i]*b)/sigma
  }
  evec[1:N] = exp_s
  evec[(N+1):(2*N)] = exp_e
  evec[(2*N+1):(3*N)] = exp_u
  return(evec)
}
mvec = rep(0,3)
mstep = function(vec, x){
  mvec[1] = mean(vec[1:N])
  mvec[2] = mean(vec[(N+1):(2*N)])
  sumx = 0; sume = 0
  for (i in 1:nrow(rikz)) {
    sumx = sumx + (x[i]^2)
    sume = sume + x[i]*vec[2*N+i]
  mvec[3] = sume/sumx
  }
  return(mvec)
}

iter = 100
for (i in 1:iter){
  e = estep(X,Y,Z,beta,D,sig)
  p = mstep(e,X)
}
p  ## output
```

The estimates of D, sigma^2, and beta are 11.3333741, 7.7893444, -0.6510261.


## Question 3
```{r}
data <- read.csv("/Users/michelle/coursework/CS/data.csv", header = TRUE)
v = 6
iter = 10  # get result of 100 patients: iter = 100
# Calculate utility function with trapezoidal, simpson's rule
utility1 = rep(0,iter)
utility2 = rep(0,iter)
for (i in 1:iter){
  X = data[i,4]
  W = data[i,1]
  A = data[i,2]
  utility1[i] = 0.5*u1trapezoid(X, W, A, 0, v/2, 1000) + 5*u2trapezoid(X, W, A, 0, v/2, 1000) + 
  10*u1trapezoid(X, W, A, v/2, v, 1000) + 20*u2trapezoid(X, W, A, v/2, v, 1000)
  utility2[i] = 0.5*u1simpson(X, W, A, 0, v/2, 1000) + 5*u2simpson(X, W, A, 0, v/2, 1000) + 
  10*u1simpson(X, W, A, v/2, v, 1000) + 20*u2simpson(X, W, A, v/2, v, 1000)
}
utility1
utility2
```

For the first 10 patients in the clinical data (data.csv), the utility function 
calculated with numerical integration using the trapezoidal rule is shown in the 
first line of the code output.  
The utility function calculated with numerical integration using simpson's 
rule is shown in the second line of the code output.  
The result of the first 100 patients can also be calculated with the code above 
(if 'iter' is changed to iter = 100), but it takes more time and the output 
becomes much longer, so only 10 patients are shown above.  


## Question 4
```{r}
# data import & preprocessing
cancer = read.table("/Users/michelle/coursework/CS/UScancer.txt", colClasses = "character")
cancer = as.matrix(cancer)
cancer = strsplit(cancer, NULL)
cancer = unlist(cancer)
cancer = as.numeric(cancer)
cancer = matrix(cancer, nrow = 58, byrow = T)
cancer[cancer==0] <- -1
cancer[cancer==2] <- 0

# estimates
mple = c(-0.3205, 0.1115)
dmh = c(-0.3020, 0.1227)
aex = c(-0.3017, 0.1224)

# Bootstrap
iter = 10
table1 = cancer
bias1 = 0
s1sum1 = rep(0,iter)
s2sum1 = rep(0,iter)
s2sum11 = rep(0,iter)
for (i in 1:iter){
  table1 = bootstrap(table1, mple[1], mple[2])
  bias1 = bias1 + (sum(table1) - sum(cancer))^2
  s1sum1[i] = sum(table1)
  s2sum1[i] = (rmse_s2(cancer, table1)[2] - rmse_s2(cancer, table1)[1])^2
  s2sum11[i] = rmse_s2(cancer, table1)[2]
}
var1 = sum((s1sum1 - mean(s1sum1))^2) / iter
bias11 = sum(s2sum1) / iter
var11 = sum((s2sum11 - mean(s2sum11))^2) / iter
rmse_s1_mple = sqrt(bias1+var1)
rmse_s2_mple = sqrt(bias11+var11)

table2 = cancer
bias2 = 0
s1sum2 = rep(0,iter)
s2sum2 = rep(0,iter)
s2sum22 = rep(0,iter)
for (i in 1:iter){
  table2 = bootstrap(table2, dmh[1], dmh[2])
  bias2 = bias2 + (sum(table2) - sum(cancer))^2
  s1sum2[i] = sum(table2)
  s2sum2[i] = (rmse_s2(cancer, table2)[2] - rmse_s2(cancer, table2)[1])^2
  s2sum22[i] = rmse_s2(cancer, table2)[2]
}
var2 = sum((s1sum2 - mean(s1sum2))^2) / iter
bias22 = sum(s2sum2) / iter
var22 = sum((s2sum22 - mean(s2sum22))^2) / iter
rmse_s1_dmh = sqrt(bias2+var2)
rmse_s2_dmh = sqrt(bias22+var22)

table3 = cancer
bias3 = 0
s1sum3 = rep(0,iter)
s2sum3 = rep(0,iter)
s2sum33 = rep(0,iter)
for (i in 1:iter){
  table3 = bootstrap(table3, aex[1], aex[2])
  bias3 = bias3 + (sum(table3) - sum(cancer))^2
  s1sum3[i] = sum(table3)
  s2sum3[i] = (rmse_s2(cancer, table3)[2] - rmse_s2(cancer, table3)[1])^2
  s2sum33[i] = rmse_s2(cancer, table3)[2]
}
var3 = sum((s1sum3 - mean(s1sum3))^2) / iter
bias33 = sum(s2sum3) / iter
var33 = sum((s2sum33 - mean(s2sum33))^2) / iter
rmse_s1_aex = sqrt(bias3+var3)
rmse_s2_aex = sqrt(bias33+var33)

# Calculate RMSE
rmse_s1_mple; rmse_s1_dmh; rmse_s1_aex
rmse_s2_mple; rmse_s2_dmh; rmse_s2_aex
```

The RMSE of alpha and beta is lowest when using estimation with MPLE (maximum 
pseudo-likelihood estimator). The difference between each RMSE of estimates of 
beta is smaller than alpha.

